---
title: "ShinyApp_3_Data_Management"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ShinyApp_3_Data_Management}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Data storage overview

One of the key contributions of the  [av_runShiny()](https://derekholmes0.github.io/alphavantagepf/reference/av_runShiny.html) app 
is to hide the details of Alphavantage asset-specific calling conventions. TO the degree possible, the app also caches locally 
that data, both to speed up retrieval and to minimize traffic to the API.  To generalize the analyses beyind that Alphavantage 
data, The app also allows other **user data** to be added.  Economic or sentiment data may be added, or rates or other company financial 
data.

The minimal set of data needed consists of a **time series** dataset and (for equities) **earnings** datasets.  For scalability and
speed, those files are kept in [fst](https://www.fstpackage.org/) format.  They can accessed directly (even when the 
app is running), or via helper functions described in the next section.  The files kept are 

|Filename|Location|Description
|:---------:|:------:|:-------------------|
|`avpf_px.fst`|Cache Directory|Raw and adjusted (total rtn) prices, and cash flows|
|`avpf_earn.fst`|Cache Directory|Historical earnings|
|`avpf_earnest.fst`|Cache Directory|Earnings forecasts|
|`avpf_inv.RD`|Cache Directory|Inventory (dates and latest values) file|

In addition, there is a constants file that is always kept in a system-assigned cache directory. This file (summarized by
[dump_state()](https://derekholmes0.github.io/alphavantagepf/reference/dump_state.html)) contains pointers to the
other files as well as downloaded ticker lists and cached state values.

# Adding new data

Each dataset described above has its own minimal set of columns and columns that may be zero for many cases. To ease the
burden of determining that, three helpful user functions are included.  These can be used in two ways, either to 
add new data or to download data from [AlphaVantage](https://www.alphavantage.co/).   Below is a list of data
available from the API and which is downloadable via the [av_runShiny()](https://derekholmes0.github.io/alphavantagepf/reference/av_runShiny.html)
app and the interface functions described in subsequent sections

|Data Item|Stored in App|Helper Function|Notes|
|:------------|:-:|:-----:|:----------------------|
|Equity,ETF prices|Y|[av_add_px](https://derekholmes0.github.io/alphavantagepf/reference/av_runShiny_userthings.html)||
|Equity Option Prices|N||Available per ticker using `OS` function|
|Equity,ETF dividends|Y|[av_add_px](https://derekholmes0.github.io/alphavantagepf/reference/av_runShiny_userthings.html)||
|Currency, Crypto prices|Y|[av_add_px](https://derekholmes0.github.io/alphavantagepf/reference/av_runShiny_userthings.html)|Not all Crypto pairs available|
|Equity related Indices|Y|[av_add_px](https://derekholmes0.github.io/alphavantagepf/reference/av_runShiny_userthings.html)|Run `AV.TICKERS` to get list[^1]|
|Equity Earnings|Y|[av_add_earn](https://derekholmes0.github.io/alphavantagepf/reference/av_runShiny_userthings.html)|Kept in `avpf_earn.fst`|
|Equity Earnings Estimates|Y|[av_add_earn](https://derekholmes0.github.io/alphavantagepf/reference/av_runShiny_userthings.html)|Kept in `avpf_earnest.fst`|
|Earnings Call Transcripts|N||Available per ticker using `EA` function|
|Company News|N||Available per ticker using `CN` function|
|Equity Financials|N||Planned for a future release|
|Insider Transactions|N||Planned for a future release|
|Commodities|N||Planned for a future release|

Any other data you may need can be added as generic (i.e. without further description) price series.

[^1]: Alphavantage has a small select list of CBOE, VIX and equity futures indices available as historical data, listed by running `AV.TICKERS`

## Adding Prices and Indices

The function [av_add_px](https://derekholmes0.github.io/alphavantagepf/reference/av_runShiny_userthings.html) can add user time series
or price data from symbols (via [av_get_pf()](https://derekholmes0.github.io/alphavantagepf/reference/av_get_pf.html)) would would normally
be downloaded from the app.

The function requires at a minimum one of two items:

* An input `data.table()` with at least three columns `c(symbol,timestamp,close)` containing the series identifier, a date, and a value.
**The user has the responsibility for ensuring that symbols are unique*.  To avoid conflicts, consider decorating your data's symbols, e.g.
`I_CL` instead of `CL` for Crude futures.  Optionally, other data (usually provided automatically from AlphaVantge) associated with intraday 
moves and total return calculations could be added.

|Data types|required?|Column names|
|:--------:|:--:|:------------------------|
|Time Series|Y|`c(symbol,timestamp,close)`|
|Intraday|N|`c(open,high,low,volume)`|
|Total Return|N|`c(adjusted_close,dividend_amount,split_coefficient)`|

Suppose we wish to download Natural Gas data from Alphavantage (via [FRED](www.stlouisfed.org) ) and give it
our own ticker `HH_GAS`. First we download the price series and get the columns we need.  Then we add some basic description,
including most critically the asset type, so the app knows where to get data going forward.

```{r, eval=FALSE}
require(data.table)
ng_dta <- av_get_pf("","NATURAL_GAS")[,.(symbol="GAS_HH",timestamp,close=value)]
asset_df <- data.frame(symbol=c("GAS_HH"),type=c("user"),currency=c("USD"), name=c("Henry Hub Gas Spot"))
av_add_px(ng_dta, assettypes=asset_df)
```
We can source data anywhere, really.  As an example of getting data directly from  [quantmod](www.quantmod.com), let's
add FEDFUNDS as its own ticker:

```{r, eval=FALSE}
suppressMessages(require(quantmod))
ffdta <- as.data.table(quantmod::getSymbols("FEDFUNDS",src="FRED",auto.assign=FALSE))
ffdta <- ffdta[,.(DT_ENTRY=index,close=FEDFUNDS,symbol="FEDFUNDS")]
av_add_px(ffdta)
```
In this case where the `assettypes` argument is not used, the source (`user`) and symbol (`symbol`) are inferred from the input data.

* A list of Equity, ETF, currency, crypto[^1] or available index[^1] symbols. For example,

```{r, eval=FALSE}
av_add_px(equitylist=c("IBM","GS","JPM"))
```

will determine the asset type, download, and inventory the data as would be done if the data were requested by a command.


## Earnings

Earnings and Earnings estimates are not strictly necessary for many of the commands, and are kept in separate files.
Like the `av_add_px()` function above, either user data can be added or a list of tickers can be given. However,
please note that **price data must always be downloaded or added before any earnings or estimates data**.

Any of the following will work:

```{r, eval=FALSE}
av_add_earn(equitylist=c("IBM","GS"))

tmp_earn <- av_get_pf("JPM","EARNINGS") |> av_extract_df("quarterlyEarnings")
tmp_earnf<- av_get_pf("JPM","EARNINGS_ESTIMATES") |> av_extract_df("estimates")
av_add_earn(substitute_earn=tmp_earn)
av_add_earn(substitute_earnest=tmp_earnf)


tmp_earn <- av_get_pf("MU","EARNINGS") |> av_extract_df("quarterlyEarnings")
tmp_earnf<- av_get_pf("MU","EARNINGS_ESTIMATES") |> av_extract_df("estimates")
av_add_earn(substitute_earn=tmp_earn, substitute_earnest=tmp_earnf)
```

```{r, echo=FALSE, eval=FALSE}
testlist <- s("GS;IBM;JPM;MU")
u1 <- lapply(testlist, \(x) kill_symbol(x))
dump_inv("^(GS|IBM|JPM|MU)")
av_add_px(equitylist = testlist)
dump_inv("^(GS|IBM|JPM|MU)")

av_add_earn(equitylist=c("IBM","GS"))
dump_inv("^(GS|IBM|JPM|MU)")

tmp_earn <- av_get_pf("JPM","EARNINGS") |> av_extract_df("quarterlyEarnings")
tmp_earnf<- av_get_pf("JPM","EARNINGS_ESTIMATES") |> av_extract_df("estimates")
av_add_earn(substitute_earn=tmp_earn)
av_add_earn(substitute_earnest=tmp_earnf)
dump_inv("^(GS|IBM|JPM|MU)")

tmp_earn <- av_get_pf("MU","EARNINGS") |> av_extract_df("quarterlyEarnings")
tmp_earnf<- av_get_pf("MU","EARNINGS_ESTIMATES") |> av_extract_df("estimates")
av_add_earn(substitute_earn=tmp_earn, substitute_earnest=tmp_earnf)
dump_inv("^(GS|IBM|JPM|MU)")
```
The advantage of such generality is that you can source price data anywhere, but not necessarily earnings data.  
Likewise, you may want to do analyses with your own forecasts, instead of consensus forecasts.

## Asset Groups

Saving sets of asset groups via the app (see [Usage](https://derekholmes0.github.io/alphavantagepf/articles/ShinyApp_1_setup_and_usage.html) is to be sure a tedious task.
To shortcut that effort, use [av_add_assetgroups()](https://derekholmes0.github.io/alphavantagepf/reference/av_add_assetgroups.html) as in the following
example:

```{r, eval=FALSE}
newtickers <- c("QQQ","QQQE","NDX")
newasset_dt <- data.table(ticker=newtickers,listnm=rep("nasdaq",length(newtickers)))
av_add_assetgroups(newasset_dt)
dump_assetgroups()
```

# Data Inventory and retrieval.

Whenever data is added, as inventory information after the addition is collected.   There are three ways to see what is currently in inventory:

* Run `AV.INV` to get all tickers with data downloaded, including indices and user data
* Run `AV.EQINV` to get just Equity and ETF tickers.
* Run [dump_inv()](https://derekholmes0.github.io/alphavantagepf/reference/av_state_interface.html) from the R console.

Also a separate tab <span style="border: 1px solid #000000; padding: 2px 6px;">INVENTORY</span> is populated on application startup.  The idea is
to always have a dictionary what what you have on hand, without going back and forth between (e.g.) `AV.INV` and your train of thought.




